---
title: "EvictionLab_Initial_Regression"
author: "Dylan Glover"
date: "July 14, 2020"
output: pdf_document
---

#Install libraries, if previously uninstalled, otherwise load them
```{r libraries, include=FALSE}
check.package.install <- function(pkge){
  is.element(pkge, installed.packages()[,1])
}

if (!check.package.install("tidyverse")){
  install.packages("tidyverse")
}

if (!check.package.install("data.table")){
  install.packages("data.table")
}

if (!check.package.install("sqldf")){
  install.packages("sqldf")
}

library(tidyverse)
library(data.table)
library(sqldf)

```

#Run if not previously downloaded EvictionLab data
#Download EvicitonLab data at tract, city levels for .csv and .geojson formats
#Set destfold variable below to your desired working directory
```{r download_files, include=FALSE}
library(USAboundaries)
state_codes

st_abbr <- state_codes$state_abbr

#Set your destination folder here
destfold <- 'C:/Users/dnsgl/Desktop/DSPG/Richmond/ELab/'

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/tracts.csv')
    str_dest <- paste0(destfold,'tracts_',st,'.csv')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/cities.csv')
    str_dest <- paste0(destfold,'cities_',st,'.csv')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/cities.geojson')
    str_dest <- paste0(destfold,'cities_',st,'.geojson')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/tracts.geojson')
    str_dest <- paste0(destfold,'tracts_',st,'.geojson')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}


```
#Upload downloaded files and combine into a dataframe: e_tracts
```{r combine_files, include=FALSE}

#tracts

e_tract_candidates = list.files(destfold,pattern="(^tracts_).*(.csv)")

setwd(destfold)

for (tract in e_tract_candidates){
       
  # Create e_tracts, if nonexistent
  if (!exists("e_tracts")){
    e_tracts <- data.table::fread(tract, header=TRUE)
  }
   
  # Append to e_tracts, if existent
  if (exists("e_tracts")){
    tmp_e_tracts <- data.table::fread(tract, header=TRUE)
    e_tracts <- rbind(e_tracts, tmp_e_tracts)
    rm(tmp_e_tracts)
  }
 
}


#cities

e_city_candidates = list.files(destfold,pattern="(^cities_).*(.csv)")

setwd(destfold)

for (city in e_city_candidates){
       
  # Create e_cities, if nonexistent
  if (!exists("e_cities")){
    e_cities <- data.table::fread(city, header=TRUE)
  }
   
  # Append to e_cities, if existent
  if (exists("e_cities")){
    tmp_e_cities <- data.table::fread(city, header=TRUE)
    e_cities <- rbind(e_cities, tmp_e_cities)
    rm(tmp_e_cities)
  }
 
}

```

```{r check_data}

library(RODBC)
con <- odbcDriverConnect("driver=SQL Server; server=LAPTOP-68S1DUHP\\WOLVERINE_NASCEN;Database=Richmond")

e_tracts$parentlocation <- e_tracts$`parent-location`
e_cities$parentlocation <- e_cities$`parent-location`

dt_tracts <- e_tracts[, names(e_tracts) := lapply(.SD, as.character)]
dt_cities <- e_cities[, names(e_cities) := lapply(.SD, as.character)]
sapply(dt_tracts, class)
sapply(dt_cities, class)

dt_tracts <- dt_tracts[,`parent-location`:=NULL]
dt_cities <- dt_cities[,`parent-location`:=NULL]

sqlSave(con, dt_tracts, tablename =
    "dt_tracts",rownames=FALSE, append = TRUE)

sqlSave(con, dt_cities, tablename =
    "dt_cities",rownames=FALSE, append = TRUE)

sqlSave(con, e_tracts, tablename =
    "R_e_tracts1",rownames=FALSE, append = TRUE, varTypes = c(rep("nvarchar(100)", ncol(e_tracts))))

sqlSave(con, e_cities, tablename =
    "R_e_cities",rownames=FALSE, append = TRUE)

```

```{r get_acs}
check.package.install <- function(pkge){
  is.element(pkge, installed.packages()[,1])
}

if (!check.package.install("tidycensus")){
  install.packages("tidycensus")
}

if (!check.package.install("tidyverse")){
  install.packages("tidyverse")
}

if (!check.package.install("stringr")){
  install.packages("stringr")
}

if (!check.package.install("ggplot2")){
  install.packages("ggplot2")
}

if (!check.package.install("gridExtra")){
  install.packages("gridExtra")
}

#Load Libraries

library(tidycensus)
library(tidyverse)
library(stringr)

# Candidate variable tables for describing employment, poverty, and educational attainment
  # B23025 
    #(001-007, 
    #EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER, 
    #Universe: EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER), 
  # B17020 
    #(001-009, 
    #POVERTY STATUS IN THE PAST 12 MONTHS BY AGE, 
    #Universe: Population For Whom Poverty Status Is Determined), 
  # B17003 
    #(001-012, 
    #POVERTY STATUS IN THE PAST 12 MONTHS OF INDIVIDUALS BY SEX BY EDUCATIONAL ATTAINMENT, 
    #Universe: Population 25 Years And Over For Whom Poverty Status Is Determined), 
  # B15003 
    #(001-025, 
    #EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER, 
    #Universe: Population 25 Years And Over), 

# Create a vector of variables to pull

vars<-c(
  tot_emp_status_16yrs_and_over_B23025_= "B23025_001", 
  emp_status_16yrs_and_over_labor_force_B23025_ = "B23025_002", 
  emp_status_16yrs_and_over_civ_labor_force_B23025_ = "B23025_003",
  emp_status_16yrs_and_over_employed_B23025_ = "B23025_004", 
  emp_status_16yrs_and_over_unemployed_B23025_ = "B23025_005", 
  emp_status_16yrs_and_over_armed_forces_B23025_ = "B23025_006", 
  emp_status_16yrs_and_over_not_in_labor_force_B23025_ = "B23025_007", 
  tot_pov_sts_12mths_age_B17020_ = "B17020_001", 
  tot_income_12mths_blw_pov_lvl_B17020_ = "B17020_002",
  income_12mths_blw_pov_lvl_under6yrs_B17020_ = "B17020_003", 
  income_12mths_blw_pov_lvl_6to11yrs_B17020_ = "B17020_004",
  income_12mths_blw_pov_lvl_12to17yrs_B17020_ = "B17020_005",
  income_12mths_blw_pov_lvl_18to59yrs_B17020_ = "B17020_006",
  income_12mths_blw_pov_lvl_60to74yrs_B17020_ = "B17020_007", 
  income_12mths_blw_pov_lvl_75to84yrs_B17020_ = "B17020_008",
  income_12mths_blw_pov_lvl_85upyrs_B17020_ = "B17020_009",
  tot_pov_sts_indiv_sex_edu_B17003_ = "B17003_001", 
  tot_income_blw_pov_pov_sts_indiv_sex_edu_B17003_ = "B17003_002", 
  income_blw_pov_male_pov_sts_indiv_sex_edu_B17003_ = "B17003_003", 
  income_blw_pov_male_no_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_004", 
  income_blw_pov_male_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_005", 
  income_blw_pov_male_assoc_pov_sts_indiv_sex_edu_B17003_ = "B17003_006", 
  income_blw_pov_male_bach_pov_sts_indiv_sex_edu_B17003_ = "B17003_007",
  income_blw_pov_female_pov_sts_indiv_sex_edu_B17003_ = "B17003_008", 
  income_blw_pov_female_no_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_009", 
  income_blw_pov_female_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_010", 
  income_blw_pov_female_assoc_pov_sts_indiv_sex_edu_B17003_ = "B17003_011", 
  income_blw_pov_female_bach_pov_sts_indiv_sex_edu_B17003_ = "B17003_012",
  tot_edu_attain_25yrs_and_up_B15003_= "B15003_001", 
  tot_edu_attain_25yrs_and_up_no_school_B15003_= "B15003_002", 
  tot_edu_attain_25yrs_and_nursery_B15003_= "B15003_003", 
  tot_edu_attain_25yrs_and_up_kindergarten_B15003_= "B15003_004", 
  tot_edu_attain_25yrs_and_up_1stgrade_B15003_= "B15003_005", 
  tot_edu_attain_25yrs_and_up_2ndgrade_B15003_= "B15003_006", 
  tot_edu_attain_25yrs_and_up_3rdgrade_B15003_= "B15003_007", 
  tot_edu_attain_25yrs_and_up_4thgrade_B15003_= "B15003_008", 
  tot_edu_attain_25yrs_and_up_5thgrade_B15003_= "B15003_009", 
  tot_edu_attain_25yrs_and_up_6thgrade_B15003_= "B15003_010",
  tot_edu_attain_25yrs_and_up_7thgrade_B15003_= "B15003_011", 
  tot_edu_attain_25yrs_and_up_8thgrade_B15003_= "B15003_012", 
  tot_edu_attain_25yrs_and_up_9thgrade_B15003_= "B15003_013", 
  tot_edu_attain_25yrs_and_up_10thgrade_B15003_= "B15003_014", 
  tot_edu_attain_25yrs_and_up_11thgrade_B15003_= "B15003_015", 
  tot_edu_attain_25yrs_and_up_12thgrade_B15003_ = "B15003_016", 
  tot_edu_attain_25yrs_and_up_HS_diploma_B15003_ = "B15003_017", 
  tot_edu_attain_25yrs_and_up_GED_B15003_ = "B15003_018", 
  tot_edu_attain_25yrs_and_up_somecollege_freshman_B15003_ = "B15003_019", 
  tot_edu_attain_25yrs_and_up_somecollege_abovefreshman_B15003_ = "B15003_020", 
  tot_edu_attain_25yrs_and_up_assoc_B15003_= "B15003_021",
  tot_edu_attain_25yrs_and_up_bach_B15003_= "B15003_022", 
  tot_edu_attain_25yrs_and_up_masters_B15003_ = "B15003_023", 
  tot_edu_attain_25yrs_and_up_prof_degree_B15003_ = "B15003_024", 
  tot_edu_attain_25yrs_and_up_doctorate_B15003_ = "B15003_025"
  )

# "get_acs" creates a census api call using the vector of variables specified above

#Apply for census api and set 'yourkey' string to your API key
yourkey <- ''

acs<-get_acs(geography = "tract",
             state="VA",
             county = "Richmond city",
             variables = vars,
             survey = "acs5",
             key = yourkey,
             output = "wide",
             show_call = T,
             geometry = T,
             keep_geo_vars = T)

#Separate the NAME column into Census_tract, County, and State
colnames=c("Census_tract","County","State")
acs<-separate(acs,NAME.y, into=colnames, sep = ", ")

# Create measures based on ACS variables
acs$perc_16yrs_and_over_in_labor_force <- 
  round((acs$emp_status_16yrs_and_over_labor_force_B23025_E / acs$tot_emp_status_16yrs_and_over_B23025_E)*100,2)
acs$perc_16yrs_and_over_unemployed <- 
  round((acs$emp_status_16yrs_and_over_unemployed_B23025_E / acs$tot_emp_status_16yrs_and_over_B23025_E)*100,2)
acs$perc_of_pov_sts_known_recent_annual_income_blw_pov_lvl <- 
  round((acs$tot_income_12mths_blw_pov_lvl_B17020_E / acs$tot_pov_sts_12mths_age_B17020_E)*100,2)
acs$percw_HS_diploma <-  
  round((acs$tot_edu_attain_25yrs_and_up_HS_diploma_B15003_E / acs$tot_edu_attain_25yrs_and_up_B15003_E)*100,2)
acs$per_below_HS_diploma <- 
  round(((acs$tot_edu_attain_25yrs_and_up_B15003_E - 
            (acs$tot_edu_attain_25yrs_and_up_assoc_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_GED_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_somecollege_abovefreshman_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_somecollege_freshman_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_HS_diploma_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_bach_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_masters_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_doctorate_B15003_E))/ acs$tot_edu_attain_25yrs_and_up_B15003_E)*100,2)

acs$per_bach_and_up <- 
  round(((acs$tot_edu_attain_25yrs_and_up_bach_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_masters_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_doctorate_B15003_E)/acs$tot_pov_sts_12mths_age_B17020_E )*100,2)


```

#Earlier attempt, disregard below
```{r read_identify_data, include=FALSE}




#national dataset
evictionlab <- data.table::fread("C:\\Users\\dnsgl\\Desktop\\DSPG\\Richmond\\Evictionlab_US_all.csv")
#state dataset for validation of national dataset
evictionlab_VA_only <- data.table::fread("C:\\Users\\dnsgl\\Desktop\\DSPG\\Richmond\\VA_all.csv") 

str(evictionlab)
str(evictionlab_VA_only)
summary(evictionlab)
summary(evictionlab_VA_only)

#cities by "name" variable
e_cities <- sqldf::sqldf("select * from evictionlab as e where e.name like ('%city')", method="raw")

str(e_cities)
summary(e_cities)

#473 "cities"
e_cities$name %>% as.data.frame() %>% distinct () %>% nrow()

#validate previous method to identify cities by uploading Virginia data
e_cities$name %>% as.data.frame() %>% distinct () %>% nrow()

b <- sqldf::sqldf("Select length(GEOID) as GEOID_prof, * from evictionlab")
evictionlab$length_GEOID <- nchar(as.character(evictionlab$GEOID))

a <- sqldf::sqldf("select * from evictionlab where GEOID like ('51%')", method="raw")

```

```

```{r profile_data, include=TRUE}
str(evictionlab)
summary(evictionlab)

e_cities <- sqldf::sqldf("select * from evictionlab as e where e.name like ('%city')", method="raw")

str(e_cities)
summary(e_cities)

#473 "cities"
e_cities$name %>% as.data.frame() %>% distinct () %>% nrow()

```
