---
title: "Create Modeling Data Source"
author: "Dylan Glover"
date: "July 27, 2020"
output: pdf_document
---

---
title: "EvictionLab_Initial_Regression"
author: "Dylan Glover"
date: "July 14, 2020"
output: pdf_document
---

#Install libraries, if previously uninstalled, otherwise load them
```{r libraries, include=FALSE}
check.package.install <- function(pkge){
  is.element(pkge, installed.packages()[,1])
}

if (!check.package.install("tidyverse")){
  install.packages("tidyverse")
}

if (!check.package.install("data.table")){
  install.packages("data.table")
}

if (!check.package.install("sqldf")){
  install.packages("sqldf")
}

if (!check.package.install("readxl")){
  install.packages("readxl")
}


if (!check.package.install("datasets.load")){
  install.packages("datasets.load")
}

if (!check.package.install("noncensus")){
  install.packages("noncensus")
}

library(tidyverse)
library(data.table)
library(sqldf)

```

#Run if not previously downloaded EvictionLab data
#Download EvicitonLab data at tract, city levels for .csv and .geojson formats
#Set destfold variable below to your desired working directory
```{r download_files, include=FALSE}
library(USAboundaries)
state_codes

st_abbr <- state_codes$state_abbr

#Use this GUI to load the counties dataset from package noncensus
datasets.load::datasets.load()

#Set your destination folder here
destfold <- 'C:/Users/dnsgl/Desktop/DSPG/Richmond/ELab/'

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/tracts.csv')
    str_dest <- paste0(destfold,'tracts_',st,'.csv')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/cities.csv')
    str_dest <- paste0(destfold,'cities_',st,'.csv')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/cities.geojson')
    str_dest <- paste0(destfold,'cities_',st,'.geojson')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}

for (st in st_abbr) {
    str <- paste0('http://eviction-lab-data-downloads.s3.amazonaws.com/',st,'/tracts.geojson')
    str_dest <- paste0(destfold,'tracts_',st,'.geojson')
   # print(str)
   # print(str_dest)
    download.file(url=str, destfile=str_dest, method="wininet", quiet=TRUE)
}


```
#Upload downloaded files and combine into a dataframe: e_tracts
```{r combine_files, include=FALSE}

#tracts

e_tract_candidates = list.files(destfold,pattern="(^tracts_).*(.csv)")

setwd(destfold)

for (tract in e_tract_candidates){
       
  # Create e_tracts, if nonexistent
  if (!exists("e_tracts")){
    e_tracts <- data.table::fread(tract, header=TRUE)
  }
   
  # Append to e_tracts, if existent
  if (exists("e_tracts")){
    tmp_e_tracts <- data.table::fread(tract, header=TRUE)
    e_tracts <- rbind(e_tracts, tmp_e_tracts)
    rm(tmp_e_tracts)
  }
 
}


#cities

e_city_candidates = list.files(destfold,pattern="(^cities_).*(.csv)")

setwd(destfold)

for (city in e_city_candidates){
       
  # Create e_cities, if nonexistent
  if (!exists("e_cities")){
    e_cities <- data.table::fread(city, header=TRUE)
  }
   
  # Append to e_cities, if existent
  if (exists("e_cities")){
    tmp_e_cities <- data.table::fread(city, header=TRUE)
    e_cities <- rbind(e_cities, tmp_e_cities)
    rm(tmp_e_cities)
  }
 
}

#relationship file

tracts_to_places <- data.table::fread('C:/Users/dnsgl/Desktop/DSPG/Richmond/tracts_to_places_geocorr2018.csv')

LT_policy <- readxl::read_excel('C:/Users/dnsgl/Desktop/DSPG/Richmond/Landlord_Tenant_Data_Policy_atlas.xlsx')

abb <- us_states()
abb <- as.data.frame(abb)

```

```{r check_data}

library(RODBC)
con <- odbcDriverConnect("driver=SQL Server; server=LAPTOP-68S1DUHP\\WOLVERINE_NASCEN;Database=Richmond")

e_tracts$parentlocation <- e_tracts$`parent-location`
e_cities$parentlocation <- e_cities$`parent-location`

tracts_to_places <- data.table::fread('C:/Users/dnsgl/Desktop/DSPG/Richmond/tracts_to_places_geocorr2018.csv')

LT_policy <- readxl::read_excel('C:/Users/dnsgl/Desktop/DSPG/Richmond/Landlord_Tenant_Data_Policy_atlas.xlsx', sheet=1)

Fair_Housing_Data <- readxl::read_excel('C:/Users/dnsgl/Desktop/DSPG/Richmond/Fair_Housing_Data.xlsx', sheet=1)

dt_tracts <- e_tracts[, names(e_tracts) := lapply(.SD, as.character)]
dt_cities <- e_cities[, names(e_cities) := lapply(.SD, as.character)]
dt_tracts_to_places <- tracts_to_places[, names(tracts_to_places) := lapply(.SD, as.character)]
sapply(dt_tracts, class)
sapply(dt_cities, class)
sapply(dt_tracts_to_places, class)

dt_tracts <- dt_tracts[,`parent-location`:=NULL]
dt_cities <- dt_cities[,`parent-location`:=NULL]

sqlSave(con, dt_tracts, tablename =
    "dt_tracts",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, dt_cities, tablename =
    "dt_cities",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, dt_tracts_to_places, tablename =
    "dt_tracts_to_places",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, e_tracts, tablename =
    "R_e_tracts1",rownames=FALSE, append = TRUE, varTypes = c(rep("nvarchar(100)", ncol(e_tracts))), fast=F)

sqlSave(con, e_cities, tablename =
    "R_e_cities",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, LT_policy, tablename =
    "LT_policy",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, abb, tablename =
    "state_meta",rownames=FALSE, append = TRUE, fast=F)

DT_LT_policy <- LT_policy %>% group_by(Jurisdiction) %>% top_n(1, Valid_Through_Date)
DT_Fair_Housing_Data <- Fair_Housing_Data %>% group_by(Jurisdictions) %>% top_n(1, Valid_Through_Date)

abb <- us_states()
abb <- as.data.frame(abb)

sqlSave(con, DT_LT_policy, tablename =
    "DT_LT_policy",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, DT_Fair_Housing_Data, tablename =
    "DT_Fair_Housing_Data",rownames=FALSE, append = TRUE, fast=F)


sqlSave(con, acs, tablename =
    "DT_acs",rownames=FALSE, append = TRUE, fast=F)


EvictionLab_Tracts_Policy <- sqlQuery(con, 'select * from dbo.EvictionLab_Tracts_Policy')

```

```{r get_acs}
check.package.install <- function(pkge){
  is.element(pkge, installed.packages()[,1])
}

if (!check.package.install("tidycensus")){
  install.packages("tidycensus")
}

if (!check.package.install("tidyverse")){
  install.packages("tidyverse")
}

if (!check.package.install("stringr")){
  install.packages("stringr")
}

if (!check.package.install("ggplot2")){
  install.packages("ggplot2")
}

if (!check.package.install("gridExtra")){
  install.packages("gridExtra")
}

#Load Libraries

library(tidycensus)
library(tidyverse)
library(stringr)

# Candidate variable tables for describing employment, poverty, and educational attainment
  # B23025 
    #(001-007, 
    #EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER, 
    #Universe: EMPLOYMENT STATUS FOR THE POPULATION 16 YEARS AND OVER), 
  # B17020 
    #(001-009, 
    #POVERTY STATUS IN THE PAST 12 MONTHS BY AGE, 
    #Universe: Population For Whom Poverty Status Is Determined), 
  # B17003 
    #(001-012, 
    #POVERTY STATUS IN THE PAST 12 MONTHS OF INDIVIDUALS BY SEX BY EDUCATIONAL ATTAINMENT, 
    #Universe: Population 25 Years And Over For Whom Poverty Status Is Determined), 
  # B15003 
    #(001-025, 
    #EDUCATIONAL ATTAINMENT FOR THE POPULATION 25 YEARS AND OVER, 
    #Universe: Population 25 Years And Over), 

# Create a vector of variables to pull

vars<-c(
  tot_emp_status_16yrs_and_over_B23025_= "B23025_001", 
  emp_status_16yrs_and_over_labor_force_B23025_ = "B23025_002", 
  emp_status_16yrs_and_over_civ_labor_force_B23025_ = "B23025_003",
  emp_status_16yrs_and_over_employed_B23025_ = "B23025_004", 
  emp_status_16yrs_and_over_unemployed_B23025_ = "B23025_005", 
  emp_status_16yrs_and_over_armed_forces_B23025_ = "B23025_006", 
  emp_status_16yrs_and_over_not_in_labor_force_B23025_ = "B23025_007", 
  tot_pov_sts_12mths_age_B17020_ = "B17020_001", 
  tot_income_12mths_blw_pov_lvl_B17020_ = "B17020_002",
  income_12mths_blw_pov_lvl_under6yrs_B17020_ = "B17020_003", 
  income_12mths_blw_pov_lvl_6to11yrs_B17020_ = "B17020_004",
  income_12mths_blw_pov_lvl_12to17yrs_B17020_ = "B17020_005",
  income_12mths_blw_pov_lvl_18to59yrs_B17020_ = "B17020_006",
  income_12mths_blw_pov_lvl_60to74yrs_B17020_ = "B17020_007", 
  income_12mths_blw_pov_lvl_75to84yrs_B17020_ = "B17020_008",
  income_12mths_blw_pov_lvl_85upyrs_B17020_ = "B17020_009",
  tot_pov_sts_indiv_sex_edu_B17003_ = "B17003_001", 
  tot_income_blw_pov_pov_sts_indiv_sex_edu_B17003_ = "B17003_002", 
  income_blw_pov_male_pov_sts_indiv_sex_edu_B17003_ = "B17003_003", 
  income_blw_pov_male_no_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_004", 
  income_blw_pov_male_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_005", 
  income_blw_pov_male_assoc_pov_sts_indiv_sex_edu_B17003_ = "B17003_006", 
  income_blw_pov_male_bach_pov_sts_indiv_sex_edu_B17003_ = "B17003_007",
  income_blw_pov_female_pov_sts_indiv_sex_edu_B17003_ = "B17003_008", 
  income_blw_pov_female_no_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_009", 
  income_blw_pov_female_hs_pov_sts_indiv_sex_edu_B17003_ = "B17003_010", 
  income_blw_pov_female_assoc_pov_sts_indiv_sex_edu_B17003_ = "B17003_011", 
  income_blw_pov_female_bach_pov_sts_indiv_sex_edu_B17003_ = "B17003_012",
  tot_edu_attain_25yrs_and_up_B15003_= "B15003_001", 
  tot_edu_attain_25yrs_and_up_no_school_B15003_= "B15003_002", 
  tot_edu_attain_25yrs_and_nursery_B15003_= "B15003_003", 
  tot_edu_attain_25yrs_and_up_kindergarten_B15003_= "B15003_004", 
  tot_edu_attain_25yrs_and_up_1stgrade_B15003_= "B15003_005", 
  tot_edu_attain_25yrs_and_up_2ndgrade_B15003_= "B15003_006", 
  tot_edu_attain_25yrs_and_up_3rdgrade_B15003_= "B15003_007", 
  tot_edu_attain_25yrs_and_up_4thgrade_B15003_= "B15003_008", 
  tot_edu_attain_25yrs_and_up_5thgrade_B15003_= "B15003_009", 
  tot_edu_attain_25yrs_and_up_6thgrade_B15003_= "B15003_010",
  tot_edu_attain_25yrs_and_up_7thgrade_B15003_= "B15003_011", 
  tot_edu_attain_25yrs_and_up_8thgrade_B15003_= "B15003_012", 
  tot_edu_attain_25yrs_and_up_9thgrade_B15003_= "B15003_013", 
  tot_edu_attain_25yrs_and_up_10thgrade_B15003_= "B15003_014", 
  tot_edu_attain_25yrs_and_up_11thgrade_B15003_= "B15003_015", 
  tot_edu_attain_25yrs_and_up_12thgrade_B15003_ = "B15003_016", 
  tot_edu_attain_25yrs_and_up_HS_diploma_B15003_ = "B15003_017", 
  tot_edu_attain_25yrs_and_up_GED_B15003_ = "B15003_018", 
  tot_edu_attain_25yrs_and_up_somecollege_freshman_B15003_ = "B15003_019", 
  tot_edu_attain_25yrs_and_up_somecollege_abovefreshman_B15003_ = "B15003_020", 
  tot_edu_attain_25yrs_and_up_assoc_B15003_= "B15003_021",
  tot_edu_attain_25yrs_and_up_bach_B15003_= "B15003_022", 
  tot_edu_attain_25yrs_and_up_masters_B15003_ = "B15003_023", 
  tot_edu_attain_25yrs_and_up_prof_degree_B15003_ = "B15003_024", 
  tot_edu_attain_25yrs_and_up_doctorate_B15003_ = "B15003_025"
  )

# "get_acs" creates a census api call using the vector of variables specified above

#create initial acs data for rbind action below
acs<-get_acs(geography = "tract",
             state="VA",
             county = "Richmond city",
             variables = vars,
             survey = "acs5",
             key = yourkey,
             output = "wide",
             geometry = T,
             keep_geo_vars = T)

#Apply for census api and set 'yourkey' string to your API key
yourkey <- ''

for (st in st_abbr) {
nn <- counties %>% as.data.frame() %>% dplyr::filter(state == st) %>% dplyr::select("county_fips")
nm <- nn$county_fips %>% as.character()

tryCatch({
acs <- rbind(acs, tidycensus::get_acs(geography = "tract",
                                      state=st,
                                      county = nm,
                                      variables = vars,
                                      survey = "acs5",
                                      key = yourkey,
                                      output = "wide",
                                      geometry = T,
                                      keep_geo_vars = T))
}, error=function(e){})

      }


acs1 <- tidycensus::get_acs(geography = "tract",
                                      state="MT",
                                      county = 'San Diego',
                                      variables = vars,
                                      survey = "acs5",
                                      key = yourkey,
                                      output = "wide",
                                      geometry = T,
                                      keep_geo_vars = T)

#Separate the NAME column into Census_tract, County, and State
colnames=c("Census_tract","County","State")
acs<-separate(acs,NAME.y, into=colnames, sep = ", ")

# Create measures based on ACS variables
acs$perc_16yrs_and_over_unemployed <- 
  round((acs$emp_status_16yrs_and_over_unemployed_B23025_E / acs$tot_emp_status_16yrs_and_over_B23025_E)*100,2)

acs$percw_HS_diploma <-  
  round((acs$tot_edu_attain_25yrs_and_up_HS_diploma_B15003_E / acs$tot_edu_attain_25yrs_and_up_B15003_E)*100,2)

acs$per_below_HS_diploma <- 
  round(((acs$tot_edu_attain_25yrs_and_up_B15003_E - 
            (acs$tot_edu_attain_25yrs_and_up_assoc_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_GED_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_somecollege_abovefreshman_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_somecollege_freshman_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_HS_diploma_B15003_E +
            acs$tot_edu_attain_25yrs_and_up_bach_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_masters_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_doctorate_B15003_E))/ acs$tot_edu_attain_25yrs_and_up_B15003_E)*100,2)


acs$per_bach_and_up <- 
  round(((acs$tot_edu_attain_25yrs_and_up_bach_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_masters_B15003_E + 
            acs$tot_edu_attain_25yrs_and_up_doctorate_B15003_E)/acs$tot_edu_attain_25yrs_and_up_B15003_E )*100,2)


```

```{r create_final_data_source}
EvictionLab_Tracts_Policy$GEOID <- as.character(EvictionLab_Tracts_Policy$GEOID)

EvictionLab_ACS_Data <- dplyr::left_join(EvictionLab_Tracts_Policy, acs, by='GEOID', suffix = c('e','acs'))

EvictionLab_ACS_Data <- EvictionLab_ACS_Data %>% as.data.table()

EvictionLab_ACS_Data <- EvictionLab_ACS_Data[, names(EvictionLab_ACS_Data) := lapply(.SD, as.character)]

data.table::fwrite(EvictionLab_ACS_Data, 'C:/Users/dnsgl/Desktop/DSPG/Richmond/github/EvictionLab_ACS_Data.csv', sep=',')

acs2 <- acs
acs2 <- as.data.table(acs2)
names(acs2) := lapply(.SD, as.character)
write.table(acs2, 'acs2.csv')


acs4 <- read.table('C:/Users/dnsgl/Desktop/DSPG/Richmond/github/dspg2020Richmond/data/acs3.txt', sep=";")


sqlSave(con, acs, tablename =
    "DT_acs",rownames=FALSE, append = TRUE, fast=F)

acs3 <- acs[,c(1:10,118:121)]
acs3 <- as.data.frame(acs3)

acs3 <- as.data.table(acs3)
acs3 <- acs3[,c(1:14)]
names(acs3) := lapply(.SD, as.character)
sqlSave(con, acs3, tablename =
    "DT_acs",rownames=FALSE, append = TRUE, fast=F)

data.table::fwrite(acs3, 'C:/Users/dnsgl/Desktop/DSPG/Richmond/github/dspg2020Richmond/data/acs3.txt', sep="")

write.table(acs3, 'C:/Users/dnsgl/Desktop/DSPG/Richmond/github/dspg2020Richmond/data/acs3.txt', sep=";")

acs3 <- read.table('C:/Users/dnsgl/Desktop/DSPG/Richmond/github/dspg2020Richmond/data/acs3.txt', sep=";")

acs3 <- acs[,c(1:6, 118:121)] %>% as.data.frame
acs3 <- acs3[,c(1:10)]
acs3 <- acs3[,c(1:10)]
acs4 <- acs4[,c(1:3,5:6,11:14)]
acs4 <- acs4[,c(1:4, 6:9)]
acs4 <- acs4[,c(4:8)]

acs4 <- do.call(data.frame,lapply(acs4, function(x) replace(x, is.infinite(x),NA)))

acs3 <- do.call(data.frame,lapply(acs3, function(x) replace(x, is.infinite(x),NA)))

sqlSave(con, acs4, tablename =
    "acs4",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, acs3, tablename =
    "acs312",rownames=FALSE, append = TRUE, fast=F)

sqlSave(con, acs5, tablename =
    "acs5",rownames=FALSE, append = TRUE, fast=F)

acs5 <- acs[,c(1:5,118:121)]
acs5 <- as.data.frame(acs5)
acs5 <- acs5[,c(1:9)]

install.packages("Amelia")
library(Amelia)


EvictionLab_Tracts_Policy2 <- sqlQuery(con, 'select * from dbo.EvictionLab_Tracts_Policy3')

EvictionLab_ACS_Data_backup <- EvictionLab_ACS_Data
EvictionLab_ACS_Data <- EvictionLab_Tracts_Policy2

E

data.table::fwrite(EvictionLab_ACS_Data %>% as.data.table(), 'C:/Users/dnsgl/Desktop/DSPG/Richmond/github/dspg2020Richmond/data/EvictionLab_ACS_Data.csv')

```


