---
title: "EvictionLab_Initial_Regression"
author: "Dylan Glover"
date: "July 27, 2020"
output: pdf_document
---

```{r, setup, include=FALSE}
#knitr::opts_knit$set(root.dir = 'C:/Users/dnsgl/Desktop/DSPG/Richmond/ELab/')

#load("EvictionLab_ACS_Data.RData")
EvictionLab_ACS_Data <- data.table::fread('EvictionLab_ACS_Data.csv')

EvictionLab_ACS_Data <- data.table::fread('C:/Users/dnsgl/Desktop/DSPG/Richmond/github/EvictionLab_ACS_Data.csv')

```

```{r create_model, include=TRUE}

set.seed(123)

#install.packages("splitstackshape")
library(splitstackshape)

#Create factor variables for regression analysis below
EvictionLab_ACS_Data$ttp_Statecode <- as.factor(EvictionLab_ACS_Data$ttp_Statecode)
EvictionLab_ACS_Data$ttp_Placecode <- as.factor(EvictionLab_ACS_Data$ttp_Placecode)
EvictionLab_ACS_Data$ttp_CountyCode <- as.factor(EvictionLab_ACS_Data$ttp_CountyCode)
EvictionLab_ACS_Data$COUSUB <- as.factor(EvictionLab_ACS_Data$COUSUB)
EvictionLab_ACS_Data$tract_name <- as.factor(EvictionLab_ACS_Data$tract_name)

#stratified sampling by state
evict_sample <- stratified(EvictionLab_ACS_Data %>% filter(city_population >= 100000), c("ttp_Statecode"), 50)

mod1 <- lm(data=evict_sample, evictionrate ~ povertyrate + 
                        medianpropertyvalue
                      + pctrenteroccupied + pcthispanic + pctasian + pctother + pctafam + 
                      pctnhpi + pctmultiple + 
                      FHP_ExcemptHousing_TOTAL + SLT_EXEMPT_TOTAL 
                      + SLT_TREMEDIES_TOTAL + SLT_SecurityDeposit + rentburden + 
                      SLT_Retaliation + 
                        SLT_LREMEDIES_TOTAL + ttp_Stateabbreviation + city  + perc_16yrs_and_over_unemployed + percw_HS_diploma + per_below_HS_diploma + per_bach_and_up)

#evictionrate ~ povertyrate + 
#                        medianpropertyvalue
 #                     + pctrenteroccupied + 
 #                     FHP_ExcemptHousing_TOTAL + SLT_EXEMPT_TOTAL 
 #                     + SLT_TREMEDIES_TOTAL + SLT_SecurityDeposit + rentburden + 
 #                       SLT_LREMEDIES_TOTAL + perc_16yrs_and_over_unemployed + #percw_HS_diploma + per_below_HS_diploma + per_bach_and_up

mod2 <- lm(data=evict_sample, evictionrate ~ povertyrate + 
                        medianpropertyvalue
                      + pctrenteroccupied + pcthispanic + pctasian + pctother + pctafam + 
                      pctnhpi + pctmultiple + 
                      FHP_ExcemptHousing_TOTAL + SLT_EXEMPT_TOTAL 
                      + SLT_TREMEDIES_TOTAL + SLT_SecurityDeposit + rentburden + 
                      SLT_Retaliation + 
                        SLT_LREMEDIES_TOTAL + perc_16yrs_and_over_unemployed + percw_HS_diploma + per_below_HS_diploma + per_bach_and_up)

mod3 <- lm(data=evict_sample, evictionrate ~ povertyrate + 
                        medianpropertyvalue
                      + pctrenteroccupied + pcthispanic + pctasian + pctother + pctafam + 
                      pctnhpi + pctmultiple + 
                      FHP_ExcemptHousing_TOTAL + SLT_EXEMPT_TOTAL 
                      + SLT_TREMEDIES_TOTAL + SLT_SecurityDeposit + rentburden + 
                      SLT_Retaliation + 
                        SLT_LREMEDIES_TOTAL + perc_16yrs_and_over_unemployed + percw_HS_diploma + per_below_HS_diploma + ttp_Stateabbreviation + city)

vif_mod2 <- car::vif(mod2) %>% as.data.frame()
colnames(vif_mod2) <- "VIF"

#evictionrate ~ 
 #                    
                        ttp_Stateabbreviation + city  + #perc_16yrs_and_over_unemployed + percw_HS_diploma + per_below_HS_diploma + per_bach_and_up



summary(mod1)

#stepwise regression


```

```{r plot}
```{r check_data_multicol, fig.height=9, fig.width=10, fig.align="center"}

es <- evict_sample[,c(13:32,37:42, 211:214)]
numericY <- unlist(lapply(es, is.numeric)) 

numerictrain <- es[,..numericY]
#initial multicollinearity plot
sj_cor = cor(numerictrain, method=c("pearson"), use="complete.obs")

#sj_cor = cor(raw_sj_train[,-c(1,3:13,15,16,19,21,24,25,26,34,28)], method=c("pearson"), use="complete.obs")

#corrplot(sj_cor, addCoef.col = "black")
 sj_cor[abs(sj_cor) > 0.6]
colnames(sj_cor) <- c(1:ncol(sj_cor))
rownames(sj_cor) <- c(1:nrow(sj_cor))

library(reshape2)
subset(melt(sj_cor), value > .60 & value < 1)


#Create datasets and apply transformations
raw_sj_train2 <- raw_sj_train[,-c(3,4,6:13,15,16,19,21,24,25,26,34,28)]

#Check missing data
raw_sj_train2[which(! complete.cases(raw_sj_train2)),]
#Assume missing at random without other information available/methods to 
#handle missing not at random

raw_sj_train3 <- raw_sj_train2[-which(! complete.cases(raw_sj_train2)),]

#final multicollinearity plot
sj_cor2 = cor(raw_sj_train[,-c(1,3:13,15,16,19,21,24,25,26,34,28)], method=c("pearson"), use="complete.obs")
corrplot(sj_cor2, addCoef.col = "black")



formula <- as.formula(evictionrate ~ povertyrate + 
                        medianpropertyvalue +
                       pctrenteroccupied + pcthispanic + pctasian + pctother + pctafam + 
                      pctnhpi + pctmultiple + 
                      FHP_ExcemptHousing_TOTAL + SLT_EXEMPT_TOTAL +
                      SLT_TREMEDIES_TOTAL + SLT_SecurityDeposit + rentburden + 
                      SLT_Retaliation + 
                        SLT_LREMEDIES_TOTAL + ttp_Stateabbreviation + city  + perc_16yrs_and_over_unemployed + percw_HS_diploma + per_below_HS_diploma + per_bach_and_up, data = evict_sample)
fit <-lm(formula)

#the linearly dependent variables
ld.vars <- attributes(alias(fit)$Complete)$dimnames[[1]]

#remove the linearly dependent variables variables
formula.new <- as.formula(
    paste(
        paste(deparse(formula), collapse=""), 
        paste(ld.vars, collapse="-"),
        sep="-"
    )
)

#run model again
fit.new <-lm(formula.new)
vif(fit.new)
```

```

